---
title: Livestock Breeding and Genomics - Notebook 2
author: Peter von Rohr
date: '2022-09-29'
output: html_notebook
params:
  doctype:
    label: Document Type
    value: solution
    choices:
    - exercise
    - solution
    - notebook
  isonline:
    label: Online (y/n)
    value: true
    choices:
    - true
    - false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Vectors in R

### Vector Definition
Although there exists a function called `vector()` in R, vectors are always defined in R using the function `c()` which stands for "concatenation". 

### Vector Assignment
Let us assume we want to assign the following vector $a$ 

$$a = \left[\begin{array}{c}
10 \\
7 \\
43
\end{array}\right]$$
to the variable named `a` in R, then this can be done with the following statement

```{r vec-assign}
a <- c(10,7,43)
```

### Access of single Vector Element
A single vector element can be accessed using the variable name followed by the element index in brackets. Hence, if we want to know the first element of vector `a`, we have to write

```{r first-element}
a[1]
```

### Computations with Vector Elements
Vector elements can be used in arithmetic operations such as summation, subtraction and multiplication as shown below

```{r arithmetic-operation}
a[1] + a[3]
a[2] * a[3]
a[3] - a[1]
```

The function `sum()` can be used to compute the sum of all vector elements. The function `mean()` computes the mean of all vector elements. 

```{r sum-mean}
sum(a)
mean(a)
```

### Vector Computations
Arithmetic operations can also be performed not only on elements of vectors but also on complete vectors. Hence, we can add the vector `a` to itself or we can multiply it by a factor of 3.5 which is shown in the following code-chunk

```{r vector-computation}
a + a
3.5 * a
```

### More Computations on Vectors

Given are the following two vectors $v$ and $w$. 

$$v = \left[\begin{array}{c}
   3  \\ 
  -5  \\ 
   1  \\ 
   9  \\
  \end{array}\right]$$
  
$$w = \left[\begin{array}{c}
    1  \\ 
    9  \\ 
  -12  \\ 
   27  \\
  \end{array}\right]$$


Compute

* the sum $v+w$, 
* the difference $v-w$ and 
* the dot product $v\cdot w$. 


### Your Solution
The following steps could be helpful for the solution

* Start by assigning the vectors to variables
* Perform the arithmetic operations with the variables
* Compute the dot-product either with `crossprod()` or with the operator `%*%`





## Problem 2: Matrices in R
Matrices in R are defined using the function `matrix()`. The function `matrix()` takes as first arguments all the elements of the matrix as a vector and as further arguments the number of rows and the number of columns. The following statment generates a matrix with $4$ rows and $3$ columns containing all integer numbers from $1$ to $12$. 

```{r}
mat_by_col <- matrix(1:12, nrow = 4, ncol = 3)
mat_by_col
```

As can be seen, the matrix elements are ordered by columns. Often, we want to define a matrix where elements are filled by rows. This can by done using the option `byrow=TRUE` 

```{r}
mat_by_row <- matrix(1:12, nrow = 4, ncol = 3, byrow = TRUE)
mat_by_row
```

### Access of Matrix Elements
Matrix elements can be accessed similarly to what was shown for vectors. But to access a single element, we need two indices, one for rows and one for columns. Hence the matrix element in the second row and third column can be accessed by 

```{r}
mat_by_row[2,3]
```


### Arithmetic Computations with Matrices
Arithmetic computations with matrices can be done with the well-known operators as long as the matrices are compatible. For summation and subtraction matrices must have the same number of rows and columns. For matrix-multiplication, the number of columns of the first matrix must be equal to the number of rows of the second matrix. 

In R the arithmetic operators `+`, `-` and `*` all perform element-wise operations. The matrix multiplication can either be done using the operator `%*%` or the function `crossprod()`. It has to be noted that the statement 

```
crossprod(A, B)
```

computes the matrix-product $A^T \cdot B$ where $A^T$ stands for the transpose of matrix $A$. Hence the matrix product $A \cdot B$ would have to be computed as 

```
crossprod(t(A), B)
```


### More Examples
Given the matrices X and Y

```{r}
X <- matrix(1:15, nrow = 5, ncol = 3)
Y <- matrix(16:30, nrow = 5, ncol = 3)
```

Compute 

- $X + Y$
- $Y - X$
- multiplication of elements between $X$ and $Y$
- matrix-product $X^T \cdot Y$
- matrix-product $X^T \cdot X$
- matrix-product $Y^T \cdot Y$



### Your Solution

* Matrices must be assigned to variables as shown above
* Arithmethic operations are performed
* Matrix muliplications can be done with `crossprod`



 


```{r, echo=FALSE, results='asis'}
cat('\n---\n\n _Latest Changes: ', format(Sys.time(), '%Y-%m-%d %H:%M:%S'), ' (', Sys.info()['user'], ')_\n', sep = '')
```
 
