---
title: Livestock Breeding and Genomics - Notebook 5
author: Peter von Rohr
date: '2023-10-13'
output: html_notebook
params:
  doctype:
    label: Document Type
    value: solution
    choices:
    - exercise
    - solution
    - notebook
  isonline:
    label: Online (y/n)
    value: true
    choices:
    - true
    - false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: QTL Data
```{r p01-setup, echo=FALSE}
s_data_path <- "https://charlotte-ngs.github.io/lbgfs2023/data/p1_qtl_1_loci.csv"  
```

Estimate genotypic values $a$ and $d$ and predict breeding values for all animals using the QTL-data given under:

```{r, echo=FALSE, results='asis'}
cat(s_data_path, "\n")
```


### Your Solution

* Start by re-formatting the paternal and maternal alleles into a genotype
* Use a linear regression to estimate genotypic values
* Predict breeding values for all animals



 

## Problem 2: Increase Effects of Genotype on Phenotype
```{r p02-setup, echo=FALSE}
s_data_path <- "https://charlotte-ngs.github.io/lbgfs2023/data/p1_qtl_1_loci.csv"  
n_h2_all <- 0.45
n_ratio_qtl <- 0.5 
```

Change the phenotypic records in the above given dataset such that the QTL explains $`r n_ratio_qtl * 100`%$ of the genetic variation when a heritability of $`r n_h2_all`$ is assumed. It is assumed that the QTL acts purely additively, hence the genotypic value of the heterozygotes can be set to $d=0$. 

Show the results as a scatter plot of all phenotypic values for the QTL genotypes.

### Your Solution

* Compute the genetic variance that can be attributed to the QTL based on the phenotypic variance on the heritability and on the amount of variation explained by the QTL. 
* Read the data and convert the paternal and the maternal alleles to QTL-genotypes
* Compute allele frequencies
* Compute the genotypic value $a$ from the QTL variance
* Add genotypic value to the phenotypes according to the QTL-Genotype
* Fit regression of new phenotypes on genotypes
* Show the results with plots




 


```{r, echo=FALSE, results='asis'}
cat('\n---\n\n _Latest Changes: ', format(Sys.time(), '%Y-%m-%d %H:%M:%S'), ' (', Sys.info()['user'], ')_\n', sep = '')
```
 
