---
title: Livestock Breeding and Genomics - Notebook 3
author: Peter von Rohr
date: '2023-09-28'
output: html_notebook
params:
  doctype:
    label: Document Type
    value: solution
    choices:
    - exercise
    - solution
    - notebook
  isonline:
    label: Online (y/n)
    value: true
    choices:
    - true
    - false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Genotype and Allele Frequencies
```{r lbg-ex03-p01-setup, echo=FALSE}
s_lbg_ex03_p01_data_url <- "https://charlotte-ngs.github.io/lbgfs2023/data/lbgfs2023_lbg_ex03.csv"
```

Given the dataset available from:

```{r, echo=FALSE}
cat(s_lbg_ex03_p01_data_url, "\n")
```

The dataset can be read using the following command

```{r, echo=FALSE, results='asis'}
cat("```r\n")
cat(paste0("readr::read_delim(\"", 
           s_lbg_ex03_p01_data_url, 
           "\", ", "\n", "delim = \",\")", 
           collapse = ""), "\n")
cat("```\n")
```


In the above dataset, genotypes are encoded as follows

```{r, echo=FALSE, message=FALSE, warning=FALSE}
tbl_genotype_code <- tibble::tibble(
  LocusG = c("$G_2G_2$",
  "$G_1G_2$",
  "$G_1G_1$"),  
  LocusH = c("$H_2H_2$",
  "$H_1H_2$",
  "$H_1H_1$"),
  Code = c(0:2)
)
knitr::kable(tbl_genotype_code, 
             booktabs = TRUE, 
             longtable = TRUE, 
             escape = FALSE)
```

### Your Tasks

* compute genotype frequencies
* compute allele frequencies

### Your Solution

* Read the dataset and assign it to a tibble or dataframe
* Compute genotype frequencies using either the function `table()` or a `dplyr` pipeline
* From genotype frequencies compute allele frequencies





## Problem 2: Check for Hardy-Weinberg Equilibrium
Use the dataset from Problem 1 and check for Hardy-Weinberg equilibrium at both loci using a $\chi^2$ test.


### Your Solution

* Read the data
* Compute genotype frequencies according to what would be expected following the Hary-Weinberg equilibrium
* Use the function `chisq.test()` for the doing the test.






```{r, echo=FALSE, results='asis'}
cat('\n---\n\n _Latest Changes: ', format(Sys.time(), '%Y-%m-%d %H:%M:%S'), ' (', Sys.info()['user'], ')_\n', sep = '')
```
 
