---
title: "The BLUP Is A Good Thing"
date: "Started: 2023-10-27 -- Last Compiled on: `r format(Sys.time(), '%Y-%m-%d')`"
output: 
  bookdown::html_document2: 
      number_sections: false
  bookdown::pdf_document2: 
      number_sections: false
      toc: false
bibliography: bibliography.bib 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
met <- rmdhelp::MendeleyExportToolR6$new()
```


## Disclaimer
This document collects arguments how BLUP works and why it is a good thing.


## Resources
The is a huge number of resources describing different aspects of BLUP. We start with the work of `r met$add("Robinson1991")`. 


## `r met$add("Robinson1991")`; That BLUP is a good thing: The estimation of random effects

### Abstract
BLUP is a technique for predicting genetic merits in animal breeding. These merits are called breeding values. In general, BLUP is a method to predict random effects and it can be used for many different applications. The paper by `r met$add("Robinson1991")` presents the theory of BLUP, some examples of its application and its relevance to the foundation of statistics. 

Understanding the procedure for predicting random effects should help people to understand some complicated and controversial issues about fixed and random effect models. It can also help to bridge the gap between Bayesian and Classical schools of thought.


### Introduction
BLUP stands for Best Linear Unbiased Prediction and is commonly used in animal breeding. It is a method to predict random effects. The context of BLUP is the linear model

$$y = X\beta + Zu + e$$

where $y$ is a vector of length $n$ with random observations, $\beta$ is a vector of length $p$ with fixed effects, $u$ is a vector of length $q$ with random breeding values and $e$ is a vector of length $n$ with random residuals. Design matrices $X$ and $Z$ link fixed effects and random breeding values to observations. Furthermore, expected values $E(u)$ and $E(e)$ of random effects $u$ and $e$ are set to 

$$E\left[\begin{array}{c}u \\ e \end{array}\right] = \left[\begin{array}{c}0 \\ 0 \end{array}\right]$$

\{Together with the model it follows that $E\left[y \right] = X\beta$\}.

Variance-Covariance matrices are given by 

$$Var\left[\begin{array}{c}u \\ e \end{array}\right] = \left[\begin{array}{cc} G & 0 \\ 0 & R\end{array}\right] \sigma^2$$
where $G$ and $R$ are known positive definite matrices and $\sigma^2$ is a positive number. 

\{We typically set $Var\left[u\right] = G$ and $Var\left[e\right] = R$ with $G = A\sigma_u^2$ and $R = I\sigma_e^2$. The reflects better the fact that when it comes to estimating dispersion parameters in a mixed model, there is not just one scalar unknown to be estimated. There is always a vector often referred to as $\theta$ which contains at least two components $\sigma_u^2$ and $\sigma_e^2$ which are to be estimated\}.

BLUP stands for Best Linear Unbiased Prediction. They are called predictions and not estimates, because they are used to predict realized values of random effects. BLUPs of the random variables $u$ are linear functions of the data $y$. They are unbiased, because $E\left[\hat{u}\right] = E\left[u\right]$. They are best in the sense of that they have minimum mean squarred error in the class of linear predictors.

Mathematically, BLUE of $\beta$ which are denoted as $\hat{\beta}$ and BLUP of $u$ are obtained as solutions to the mixed model equations, given by

$$X^TR^{-1}X\hat{\beta} + X^TR^{-1}Z\hat{u} = X^TR^{-1}y $$
$$Z^TR^{-1}X\hat{\beta} + (Z^TR^{-1}Z + G^{-1})\hat{u} = Z^TR^{-1}y $$

It is noted that if matrix $G^{-1}$ tends towards the zero matrix the above equations tend towards the generalized least squares normal equations for estimating $\beta$ and $u$ as fixed effects. 


```{r, echo=FALSE, results='asis'}
# put references before latest change
if (exists("met")){
  cat("# References {-}\n\n")
  if (knitr::is_html_output()){
    cat("<div id=\"refs\"></div>\n\n")
  }
}
# write latest change message
if (knitr::is_html_output() && is.element("rmdhelp", installed.packages()))
  cat(paste0(rmdhelp::get_latest_change()))
```
