---
title: "Prepare RStudio-Server Exercise Platform"
author: "Peter von Rohr"
date: "2023-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disclaimer
Preparation steps to get to the RStudio-Server platform are documented.


## Procedure
The following steps are followed to prepare required files for students.


### Step 1: Download BelegungenLerneinheit from edoz
Login to `edoz` and download the BelegungenLerneinheit as xlsx file. Move the file to a separate directory.

```bash
mkdir students
cd students/
mv ~/Downloads/BelegungenLerneinheit_751630500L_Herbstsemester_2023.xlsx .
```


### Step 2: Process Downloaded xlsx file
From the downloaded xlsx file, we only need some columns

```{r}
s_bl_fs_2023_dir <- file.path(here::here(), "students")
s_bl_fs_2023_path <- list.files(path = s_bl_fs_2023_dir, pattern = "xlsx", full.names = TRUE)
tbl_bl_fs_2023 <- readxl::read_excel(path = s_bl_fs_2023_path)
tbl_bl_fs_2023
```

Select only the columns that are required

```{r}
library(dplyr)
tbl_students_fs_2023 <- tbl_bl_fs_2023 %>%
  select(Familienname, Rufname, Nummer, `E-Mail`)
tbl_students_fs_2023
```

### Step 3: Generate Usernames
The part of the email addresses before the `@` are taken as usernames

```{r}
tbl_students_fs_2023$Username <- sapply(tbl_students_fs_2023$`E-Mail`, 
                                         function(x) unlist(strsplit(x, "@", fixed = T))[1] , 
                                         USE.NAMES = FALSE)
tbl_students_fs_2023
```


### Step 4: Adding Ports
```{r, echo=FALSE}
n_rstudio_port <- 8787
n_start_port <- 10087
```

The docker instances are mapping the RStudio-Server port `r n_rstudio_port` to different ports on the host. These are generated from `r n_start_port` onwards and directly written to the student info file.

```{r}
tbl_students_fs_2023$Port <- c(n_start_port:(n_start_port + nrow(tbl_students_fs_2023)-1))
tbl_students_fs_2023
```



### Step 5: Write Info To Files
Create two files, one with the complete student information and one with only usernames

```{r}
# student info
s_students_fs_2023_path <- file.path(s_bl_fs_2023_dir, "students_lbgfs_2023.csv")
readr::write_delim(tbl_students_fs_2023, 
                   file = s_students_fs_2023_path, 
                   delim = ",",
                   col_names = FALSE)
# usernames
tbl_usernames_fs_2023 <- tbl_students_fs_2023 %>%
  select(Username)
s_usernames_fs_2023_path <- file.path(s_bl_fs_2023_dir, "usernames_lbgfs_2023.csv")
readr::write_delim(tbl_usernames_fs_2023, 
                   file = s_usernames_fs_2023_path, 
                   delim = ",",
                   col_names = FALSE)
```


## Test Student
Do the same for a test student

```{r}
tbl_test_students_fs_2023 <- tibble::tibble(Familienname = "von Rohr",
                                            Rufname      = "Peter",
                                            Nummer       = "88-918-123",
                                            `E-Mail`     = "peter.vonrohr@usys.ethz.ch",
                                            Username     = "vrohrp",
                                            Port         = 10187)
s_test_students_fs_2023_path <- file.path(s_bl_fs_2023_dir, "test_students_lbgfs_2023.csv")
readr::write_delim(tbl_test_students_fs_2023, 
                   file = s_test_students_fs_2023_path, 
                   delim = ",",
                   col_names = FALSE)
tbl_test_usernames_fs_2023 <- tbl_test_students_fs_2023 %>%
  select(Username)
s_test_usernames_fs_2023_path <- file.path(s_bl_fs_2023_dir, "test_usernames_lbgfs_2023.csv")
readr::write_delim(tbl_test_usernames_fs_2023, 
                   file = s_test_usernames_fs_2023_path, 
                   delim = ",",
                   col_names = FALSE)

```


## Upload Data to Server
Prepare on server

```bash
# sq2h
cd;cd courses/
mkdir -p lbgfs2023
cd lbgfs2023/
```

Upload

```bash
cd lbgfs2023_gh-root/main/lbgfs2023/
scp -r docker quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/lbgfs2023
scp -r students quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/lbgfs2023
scp -r bash quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/lbgfs2023
```

Build new container image

```bash
# rm old images first
docker rmi <image_ids>
# build from dockerfile
cd;cd courses/lbgfs2023/docker
docker build -t lbgfs2023 .
```


## Start Container Instances

```bash

```


## Send E-Mail
The following commands are used to send e-mails to students

```
export GELSRC=/home/quagadmin/courses/lbgfs2023
cd $GELSRC
if [ ! -d "email_sent" ]; then mkdir -p email_sent;fi
cat students/test_students_lbgfs_2023.csv  | while read s
do
  user=$(echo "$s" | cut -d ',' -f5)
  email=$(echo "$s" | cut -d ',' -f4)
  echo " * Sending ${user}_mail.txt to $email ..."
  cat ${user}_email.txt | ssmtp $email
  sleep 2
  mv ${user}_email.txt email_sent
done

# real students
cat students/students_lbgfs_2023.csv  | while read s
do
  user=$(echo "$s" | cut -d ',' -f5)
  email=$(echo "$s" | cut -d ',' -f4)
  echo " * Sending ${user}_mail.txt to $email ..."
  cat ${user}_email.txt | ssmtp $email
  sleep 2
  mv ${user}_email.txt email_sent
done
```


## Create Table with Rexpf Links
Links are created in separate Rmd document.


## Cloning the material

```
# on 2-htz
export GITURI=https://github.com/charlotte-ngs/lbgfs2023.git
export GELSRC=/home/quagadmin/courses/lbgfs2023
export GELHOME=${GELSRC}/home
cd $GELSRC
# test student
cat students/test_usernames_lbgfs_2023.csv | while read s
do
  cd $GELHOME/$s
  echo " * Cloning repo for $s ..."
  git clone $GITURI -b rexpf
  sleep 2
done

# students
cd $GELSRC
# real student
cat students/usernames_lbgfs_2023.csv | while read s
do
  cd $GELHOME/$s
  echo " * Cloning repo for $s ..."
  git clone $GITURI -b rexpf
  sleep 2
done
```

