---
title: Livestock Breeding and Genomics - `r tools::toTitleCase(params$doctype)` 3
author: Peter von Rohr
date: 2023-09-28
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_md: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Genotype and Allele Frequencies
```{r lbg-ex03-p01-setup, echo=FALSE}
s_lbg_ex03_p01_data_url <- "https://charlotte-ngs.github.io/lbgfs2023/data/lbgfs2023_lbg_ex03.csv"
```

Given the dataset available from:

```{r, echo=FALSE}
cat(s_lbg_ex03_p01_data_url, "\n")
```

The dataset can be read using the following command

```{r, echo=FALSE, results='asis'}
cat("```r\n")
cat(paste0("readr::read_delim(\"", 
           s_lbg_ex03_p01_data_url, 
           "\", ", "\n", "delim = \",\")", 
           collapse = ""), "\n")
cat("```\n")
```


In the above dataset, genotypes are encoded as follows

```{r, echo=FALSE, message=FALSE, warning=FALSE}
tbl_genotype_code <- tibble::tibble(
  Genotype = c("$G_2G_2$",
  "$G_1G_2$",
  "$G_1G_1$"),
  Code = c(0:2)
)
knitr::kable(tbl_genotype_code, 
             booktabs = TRUE, 
             longtable = TRUE, 
             escape = FALSE)
```

### Your Tasks

* compute genotype frequencies
* compute allele frequencies

<!-- your-solution-start
### Your Solution

* Read the dataset and assign it to a tibble or dataframe
* Compute genotype frequencies using either the function `table()` or a `dplyr` pipeline
* From genotype frequencies compute allele frequencies

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

<!-- master-solution-end -->


## Problem 2: Check for Hardy-Weinberg Equilibrium
Use the dataset from Problem 1 and check for Hardy-Weinberg equilibrium using a $\chi^2$ test.


<!-- your-solution-start
### Your Solution

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

<!-- master-solution-end -->


## Problem 3

<!-- your-solution-start
### Your Solution

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

<!-- master-solution-end -->
 
