---
title: Livestock Breeding and Genomics - `r tools::toTitleCase(params$doctype)` 3
author: Peter von Rohr
date: 2023-09-28
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_md: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Genotype and Allele Frequencies
```{r lbg-ex03-p01-setup, echo=FALSE}
s_lbg_ex03_p01_data_url <- "https://charlotte-ngs.github.io/lbgfs2023/data/lbgfs2023_lbg_ex03.csv"
```

Given the dataset available from:

```{r, echo=FALSE}
cat(s_lbg_ex03_p01_data_url, "\n")
```

The dataset can be read using the following command

```{r, echo=FALSE, results='asis'}
cat("```r\n")
cat(paste0("readr::read_delim(\"", 
           s_lbg_ex03_p01_data_url, 
           "\", ", "\n", "delim = \",\")", 
           collapse = ""), "\n")
cat("```\n")
```


In the above dataset, genotypes are encoded as follows

```{r, echo=FALSE, message=FALSE, warning=FALSE}
tbl_genotype_code <- tibble::tibble(
  LocusG = c("$G_2G_2$",
  "$G_1G_2$",
  "$G_1G_1$"),  
  LocusH = c("$H_2H_2$",
  "$H_1H_2$",
  "$H_1H_1$"),
  Code = c(0:2)
)
knitr::kable(tbl_genotype_code, 
             booktabs = TRUE, 
             longtable = TRUE, 
             escape = FALSE)
```

### Your Tasks

* compute genotype frequencies
* compute allele frequencies

<!-- your-solution-start
### Your Solution

* Read the dataset and assign it to a tibble or dataframe
* Compute genotype frequencies using either the function `table()` or a `dplyr` pipeline
* From genotype frequencies compute allele frequencies

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

* Read the dataset and assign it to a tibble or dataframe

```{r, message=FALSE, warning=FALSE}
s_lbg_ex03_p01_data_url <- "https://charlotte-ngs.github.io/lbgfs2023/data/lbgfs2023_lbg_ex03.csv"
tbl_lbe_ex03 <- readr::read_delim(s_lbg_ex03_p01_data_url, delim = ",")
tbl_lbe_ex03
```


* Compute genotype frequencies using either the function `table()` or a `dplyr` pipeline
The solution with `table()` results in counts which can be converted to frequencies

```{r}
n_nr_total_animals <- nrow(tbl_lbe_ex03)
vec_freq_table_LocusG <- table(tbl_lbe_ex03$LocusG)
round(vec_freq_table_LocusG / n_nr_total_animals, digits = 3)
```

The same for locus $H$

```{r}
vec_freq_table_LocusH <- table(tbl_lbe_ex03$LocusH)
round(vec_freq_table_LocusH / n_nr_total_animals, digits = 3)
```

The second solution is to use `dplyr`

```{r, message=FALSE, warning=FALSE}
library(dplyr)
tbl_lbe_ex03 %>% 
  select(LocusG) %>%
  group_by(LocusG) %>% 
  summarise(genotype_frequency = n() / n_nr_total_animals)
```

Similarly for Locus $H$

```{r, message=FALSE, warning=FALSE}
tbl_lbe_ex03 %>% 
  select(LocusH) %>%
  group_by(LocusH) %>% 
  summarise(genotype_frequency = n() / n_nr_total_animals)
```

<!-- master-solution-end -->


## Problem 2: Check for Hardy-Weinberg Equilibrium
Use the dataset from Problem 1 and check for Hardy-Weinberg equilibrium at both loci using a $\chi^2$ test.


<!-- your-solution-start
### Your Solution

* Read the data
* Compute genotype frequencies according to what would be expected following the Hary-Weinberg equilibrium
* Use the function `chisq.test()` for the doing the test.
---  your-solution-end -->


<!-- master-solution-start -->

### Solution

* Read the data, as shown above

```{r, message=FALSE, warning=FALSE}
s_lbg_ex03_p01_data_url <- "https://charlotte-ngs.github.io/lbgfs2023/data/lbgfs2023_lbg_ex03.csv"
tbl_lbe_ex03 <- readr::read_delim(s_lbg_ex03_p01_data_url, delim = ",")
```


* Genotype frequencies: Because of the used genotype encoding, the numbers can be interpreted as counts of the favorable alleles $G_1$ and $H_1$. So the total number of favorable alleles for both loci is given by

```{r}
sum(tbl_lbe_ex03$LocusG)
```

for Locus $G$ and for Locus $H$

```{r}
sum(tbl_lbe_ex03$LocusH)
```

To get to the allele frequencies $p$ at both loci, we have to divide these sums by the total number of alleles which is two times the number of animals in the dataset. Hence the allele frequencies are

```{r}
n_nr_alleles_total <- 2 * n_nr_total_animals
round(sum(tbl_lbe_ex03$LocusG) / n_nr_alleles_total, digits = 3)
```

for Locus $G$ and

```{r}
round(sum(tbl_lbe_ex03$LocusH) / n_nr_alleles_total, digits = 3)
```

These frequencies can also be computed by taking the mean of each of the genotype columns and deviding them by $2$

```{r}
round(mean(tbl_lbe_ex03$LocusG)/2, digits = 3)
```

and analogeously

```{r}
round(mean(tbl_lbe_ex03$LocusH)/2, digits = 3)
```

Hence the allele frequencies are 

```{r}
n_allele_freq_p_locus_G <- mean(tbl_lbe_ex03$LocusG)/2
n_allele_freq_q_locus_G <- 1-n_allele_freq_p_locus_G
n_allele_freq_p_locus_H <- mean(tbl_lbe_ex03$LocusH)/2
n_allele_freq_q_locus_H <- 1-n_allele_freq_p_locus_H
```

The genotype frequencies according to Hardy-Weinberg are given by

```{r}
vec_geno_freq_locus_G_hw <- c(n_allele_freq_q_locus_G^2,
                           2*n_allele_freq_q_locus_G * n_allele_freq_p_locus_G,
                           n_allele_freq_p_locus_G^2)
round(vec_geno_freq_locus_G_hw, digits = 3)
```

and the same for locus $H$

```{r}
vec_geno_freq_locus_H_hw <- c(n_allele_freq_q_locus_H^2,
                           2*n_allele_freq_q_locus_H * n_allele_freq_p_locus_H,
                           n_allele_freq_p_locus_H^2)
round(vec_geno_freq_locus_H_hw, digits = 3)
```


* Run $\chi^2$ test with then genotype counts from the data

```{r}
chisq.test(table(tbl_lbe_ex03$LocusG), p = vec_geno_freq_locus_G_hw)
```

for locus $H$

```{r}
chisq.test(table(tbl_lbe_ex03$LocusH), p = vec_geno_freq_locus_H_hw)
```

Although there are differences in the distribution of genotype frequencies, the $\chi^2$ test does not give us a result in the test-statistic that would suggest that there is a detectable deviation from the Hary-Weinberg equilibrium.






<!-- master-solution-end -->


