---
title: Livestock Breeding and Genomics - Notebook 8
author: Peter von Rohr
date: '2023-11-03'
output: html_notebook
params:
  doctype:
    label: Document Type
    value: solution
    choices:
    - exercise
    - solution
    - notebook
  isonline:
    label: Online (y/n)
    value: true
    choices:
    - true
    - false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Regression Model
```{r l06prob01-setup, echo=FALSE}
if (params$isonline){
  s_data_path <- "https://charlotte-ngs.github.io/lbgfs2023/data/beef_data_bc.csv"
} else {
  s_data_path <- file.path(here::here(), "docs", "data", "beef_data_bc.csv")
}
```

Use the dataset on weaning weight and fit a regression model of weaning weight on breast circumference. The following tasks are to be completed.

* Use matrix-vector notation to specify the model and fill the information from the dataset into the model. 
* Compute the solution for the estimated regression coefficient using a least squares approach. 
* Use R to verify your result.

The data set is available from

```{r, echo=FALSE}
cat(s_data_path, "\n")
```


### Your Solution

* Regression model in matrix-vector notation: define the vectors and matrices required in the model
* Read the data into a tibble/data_frame
* Put information from dataset into the model
* Compute solution for regression coefficient estimate
* Use R to verify





## Problem 2: Fixed Linear Effects Model
Use the same dataset as in Problem 1 and fit a fixed linear effects model using breast circumference and herd as fixed effects in a model. Use the same path to get to the solution as in Problem 1 and complete the same set of tasks.


### Your Solution






```{r, echo=FALSE, results='asis'}
cat('\n---\n\n _Latest Changes: ', format(Sys.time(), '%Y-%m-%d %H:%M:%S'), ' (', Sys.info()['user'], ')_\n', sep = '')
```
 
