---
title: Livestock Breeding and Genomics - Notebook 6
author: Peter von Rohr
date: '2023-10-20'
output: html_notebook
params:
  doctype:
    label: Document Type
    value: solution
    choices:
    - exercise
    - solution
    - notebook
  isonline:
    label: Online (y/n)
    value: true
    choices:
    - true
    - false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Parent Offspring Breeding Values
As shown in the course notes, the breeding value $u_i$ of animal $i$ can be decomposed into the average of the parent breeding values plus a mendelian sampling term ($m_i$). This means

$$u_i = {1\over 2}u_s + {1\over 2}u_d + m_i$$

where animal $i$ has parents $s$ and $d$. The mendelian sampling term $m_i$ is the deviation of the single breeding value $u_i$ from the parent average breeding value. Because $m_i$ is modelled as a deviation, it follows that for a large number ($N$) of offspring from parents $s$ and $d$, the average over all mendelian sampling terms must be $0$. 

### Your Task
Show that the average mendelian sampling term over a large number of offspring is $0$ using a single locus model for the following cases.

#### Case 1: Homozygous and Heterozygous Parents
Parent $s$ with genotype $G_1G_1$ and parent $d$ with genotype $G_1G_2$


#### Case 2: Homozygous and Heterozygous Parents
Parent $s$ with genotype $G_2G_2$ and parent $d$ with genotype $G_1G_2$


#### Case 3: Heterozygous Parents
Both parents $s$ and $d$ have genotype $G_1G_2$

### Your Solution
* For each of the following cases compute the parent average of breeding values. 
* Compute the difference between the breeding values of every possible offspring and the parent average
* Compute the average over all mendelian sampling terms

#### Case 1: Homozygous and Heterozygous Parents

#### Case 2: Homozygous and Heterozygous Parents

#### Case 3: Heterozygous Parents







```{r, echo=FALSE, results='asis'}
cat('\n---\n\n _Latest Changes: ', format(Sys.time(), '%Y-%m-%d %H:%M:%S'), ' (', Sys.info()['user'], ')_\n', sep = '')
```
 
