---
title: Livestock Breeding and Genomics - `r tools::toTitleCase(params$doctype)` 8
author: Peter von Rohr
date: 2023-11-03
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_md: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Regression Model
```{r l06prob01-setup, echo=FALSE}
if (params$isonline){
  s_data_path <- "https://charlotte-ngs.github.io/lbgfs2023/data/beef_data_bc.csv"
} else {
  s_data_path <- file.path(here::here(), "docs", "data", "beef_data_bc.csv")
}
```

Use the dataset on weaning weight and fit a regression model of weaning weight on breast circumference. The following tasks are to be completed.

* Use matrix-vector notation to specify the model and fill the information from the dataset into the model. 
* Compute the solution for the estimated regression coefficient using a least squares approach. 
* Use R to verify your result.

The data set is available from

```{r, echo=FALSE}
cat(s_data_path, "\n")
```


<!-- your-solution-start
### Your Solution

* Regression model in matrix-vector notation: define the vectors and matrices required in the model
* Read the data into a tibble/data_frame
* Put information from dataset into the model
* Compute solution for regression coefficient estimate
* Use R to verify

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

The regression model in matrix-vector notation is given by the following formula

$$y = X\beta + e$$

where $y$ is the vector of known responses. For our example this is the vector of weaning weights. The vector $\beta$ contains the intercept and the regression coefficient. The vector $e$ contains the random residuals. The matrix $X$ contains a column of all ones and a second column with breast circumference. 

The data are imported to a tibble using 

```{r}
tbl_ww <- readr::read_delim(file = s_data_path, delim = ",",
                            col_types = readr::cols(
                              Animal = readr::col_integer(),
                              Sire = readr::col_integer(),
                              Dam = readr::col_integer(),
                              Herd = readr::col_factor(),
                              `Weaning Weight` = readr::col_double(),
                              `Breast Circumference` = readr::col_double()
                            ))
head(tbl_ww)
```

The information for the known vector $y$ and for matrix $X$ are taken from the dataset. The vector $y$ contains all weaning weight values 

```{r}
vec_y <- tbl_ww$`Weaning Weight`
```

The matrix $X$ is a matrix with two columns and as many rows as there are observations.

```{r}
n_nr_obs <- nrow(tbl_ww)
mat_X <- matrix(c(rep(1,n_nr_obs), tbl_ww$`Breast Circumference`), nrow = n_nr_obs, ncol = 2)
```

The least squares estimate is

$$\hat{\beta} = (X^TX)^{-1}X^Ty$$

Inserting the numbers into the formula leads to 

```{r}
mat_xtx <- crossprod(mat_X)
mat_xty <- crossprod(mat_X, vec_y)
vec_hat_beta <- solve(mat_xtx, mat_xty)
vec_hat_beta
```

The standard deviation of the residuals is used as an estimate of the spread of the error terms.

```{r}
vec_r <- vec_y - crossprod(t(mat_X), vec_hat_beta)
n_sd_res <- sqrt(crossprod(vec_r) / (n_nr_obs-2))
n_sd_res
```


Verification of the results using R

```{r}
lm_ww <- lm(`Weaning Weight` ~ `Breast Circumference`, data = tbl_ww)
summary(lm_ww)
```

<!-- master-solution-end -->


## Problem 2: Fixed Linear Effects Model
Use the same dataset as in Problem 1 and fit a fixed linear effects model using breast circumference and herd as fixed effects in a model. Use the same path to get to the solution as in Problem 1 and complete the same set of tasks.


<!-- your-solution-start
### Your Solution

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

The model in matrix vector notation is the same as in Problem 1

$$y = X\beta + e$$

But the definition of the vector $\beta$ and the matrix $X$ are different. The vectors $y$ and $e$ are defined the same way.

The check with R

```{r}
lm_ww_bc_herd <- lm(`Weaning Weight` ~ `Breast Circumference` + Herd, data = tbl_ww)
summary(lm_ww_bc_herd)
```

<!-- master-solution-end -->

