---
title: Livestock Breeding and Genomics - Notebook 9
author: Peter von Rohr
date: '2023-11-10'
output: html_notebook
params:
  doctype:
    label: Document Type
    value: solution
    choices:
    - exercise
    - solution
    - notebook
  isonline:
    label: Online (y/n)
    value: true
    choices:
    - true
    - false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Numerator Relationship Matrix
```{r lbg-ex07-p01-setup, echo=FALSE, message=FALSE, warning=FALSE}
s_ped_data <- "https://charlotte-ngs.github.io/lbgfs2023/data/ped_num_rel_mat.csv"
```

The following pedigree is given

```{r, echo=FALSE}
nr_animal <- 6
tbl_pedigree <- readr::read_delim(file = s_ped_data, 
                                  col_types = readr::cols(Calf = readr::col_integer(),
                                                          Sire = readr::col_integer(),
                                                          Dam = readr::col_integer()))
knitr::kable(tbl_pedigree, booktabs = TRUE, longtable = TRUE)
```

The pedigree can be read from the file

```{r, echo=FALSE}
cat(s_ped_data, "\n")
```

Compute the numerator relationship matrix $A$ for the given pedigree. Recall from the course notes that elements of matrix $A$ are computed differently for elements on the diagonal and for off-diagonal elements. In summary, we compute

* diagonal element $(A)_{ii}$ as $(A)_{ii} = 1+F_i$ where $F_i = 0.5 * (A)_{sd}$ where $s$ and $d$ are parents of $i$. 
* off-diagonal element $(A)_{ki}$ as $(A)_{ki} = 0.5 * \left[(A)_{ks} + (A)_{kd}\right]$ where $s$ and $d$ are parents of $i$


### Task
Use two nested loops over the rows and the columns of matrix $A$ to compute all the elements of matrix $A$ using the formulas given above.

### Your Solution

* Read the pedigree from the given file
* Add parents that do not appear as animals to the column of animals
* Create an empty square matrix $A$ with as many rows and columns as we have animals in the pedigree.
* Use the above formula to compute the diagonal element of the first row
* Create a first loop to compute all the off-diagonal elements of the first row
* Create a loop to repeat the computations over all rows.





## Problem 2: Verification
Use the function `pedigreemm::getA()` from package `pedigreemm` to verify your result from problem 1.

### Your Solution

* Define the pedigree
* Compute the relationship matrix with `getA()`





## Problem 3: Functions in R
Computations such as the computation of the diagonal elements or such as the one of the off-diagonal elements can be isolated and factored out in a important programming construct which is called `function`. A function takes a set of input parameter and transforms them into a result which is returned. For our example of the numerator relationship matrix two functions can be constructed according to the following template

```{r}
compute_square <- function(pn_number){
  square_result <- pn_number*pn_number
  return(square_result)
}
```

The function can be used by function calls which take a given input and return a result

```{r}
compute_square(pn_number = 3)
compute_square(1:10)
```


### Task
Use the above template to construct a function and factor out the computations of the diagonal elements and of the off-diagonal elements into two separate functions.


### Your Solution

* Write a function to compute a diagonal element
* Write a function to compute an off-diagonal element
* Use the functions to compute all elements of the numerator relationship matrix



 


```{r, echo=FALSE, results='asis'}
cat('\n---\n\n _Latest Changes: ', format(Sys.time(), '%Y-%m-%d %H:%M:%S'), ' (', Sys.info()['user'], ')_\n', sep = '')
```
 
