---
title: Livestock Breeding and Genomics - Notebook 12
author: Peter von Rohr
date: '2023-12-01'
output: html_notebook
params:
  doctype:
    label: Document Type
    value: solution
    choices:
    - exercise
    - solution
    - notebook
  isonline:
    label: Online (y/n)
    value: true
    choices:
    - true
    - false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r p1-init, echo=FALSE}
if (!exists("params")) params <- list(isonline = TRUE)
if (params$isonline){
  s_beef_data_path <- "https://charlotte-ngs.github.io/lbgfs2023/data/beef_data_blup.csv"
} else {
  s_beef_data_path <- here::here("docs", "data", "beef_data_blup.csv")
}
n_h2_weight <- 0.25
```


## Problem 1: Accuracy of Predicted Breeding Values
Use the dataset given below to predict breeding values for the response variable `Weight` using a BLUP animal model. The model contains `Herd` as fixed effect and `BreastCircumference` as regression covariate. Compute reliabilities ($B\%$) for all predicted breeding values. 


### Hints
* The phenotypic variance $\sigma_p^2$ can be computed from the variance of the weight values given in the dataset. 
* Heritability ($h^2$) is assumed to be `r n_h2_weight` for the trait `Weight`. 


### Data
The dataset is available from

```{r, echo=FALSE}
cat(s_beef_data_path)
```



### Your Solution

* Read the data
* Specify the linear mixed effects model
* Setup the mixed model equations (MME)
* Get estimates of fixed effects and predictions of breeding values from solutions of MME
* Compute accuracies from inverse of coefficient matrix of MME







```{r, echo=FALSE, results='asis'}
cat('\n---\n\n _Latest Changes: ', format(Sys.time(), '%Y-%m-%d %H:%M:%S'), ' (', Sys.info()['user'], ')_\n', sep = '')
```
 
