---
title: Livestock Breeding and Genomics - `r tools::toTitleCase(params$doctype)` 12
author: Peter von Rohr
date: 2023-12-01
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_md: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r p1-init, echo=FALSE}
if (!exists("params")) params <- list(isonline = TRUE)
if (params$isonline){
  s_beef_data_path <- "https://charlotte-ngs.github.io/lbgfs2023/data/beef_data_blup.csv"
} else {
  s_beef_data_path <- here::here("docs", "data", "beef_data_blup.csv")
}
n_h2_weight <- 0.25
```


## Problem 1: Accuracy of Predicted Breeding Values
Use the dataset given below to predict breeding values for the response variable `Weight` using a BLUP animal model. The model contains `Herd` as fixed effect and `BreastCircumference` as regression covariate. Compute reliabilities ($B\%$) for all predicted breeding values. 


### Hints
* The phenotypic variance $\sigma_p^2$ can be computed from the variance of the weight values given in the dataset. 
* Heritability ($h^2$) is assumed to be `r n_h2_weight` for the trait `Weight`. 


### Data
The dataset is available from

```{r, echo=FALSE}
cat(s_beef_data_path)
```



<!-- your-solution-start
### Your Solution

* Read the data
* Specify the linear mixed effects model
* Setup the mixed model equations (MME)
* Get estimates of fixed effects and predictions of breeding values from solutions of MME
* Compute accuracies from inverse of coefficient matrix of MME

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

* Read the data

```{r}
tbl_beef <- readr::read_delim(s_beef_data_path, delim = ",")
```

* Specify the linear mixed effects model

$$y = X\beta + Zu + e$$
with vectors $y$: observations, $\beta$: fixed regression coefficient of `BreastCircumference` and fixed herd-effects, $u$: random breeding values, $e$: random residuals. Matrix $X$ links covariates and fixed effects to observations and matrix Z related breeding values to observations. 

Expected values of the random components are 

$$E \left[\begin{array}{c}y \\ u \\ e  \end{array} \right] = \left[\begin{array}{c}X\beta \\ 0 \\ 0  \end{array} \right]$$

Variance-Covariance matrices of the random components are

$$var\left[\begin{array}{c}y \\ u \\ e  \end{array} \right] = \left[\begin{array}{ccc} V & ZG & R \\ GZ^T & G & 0 \\ R & 0 & R  \end{array} \right] $$
with $R = I * \sigma_e^2$, $G = A \sigma_u^2$ and $V = ZGZ^T + R$. Matrix $A$ is the numerator relationship matrix based on the pedigree and $\sigma_e^2$ and $\sigma_u^2$ are given variance components. 


* Setup the mixed model equations (MME)

Mixed model equations for BLUE of $\beta$ and BLUP of $u$ are given by

$$
\left[
\begin{array}{lr}
X^TX & X^TZ \\
Z^TX & Z^TZ + \lambda * A^{-1}
\end{array}
\right]
\left[
\begin{array}{c}
\hat{\beta} \\
\hat{u}
\end{array}
\right]
=
\left[
\begin{array}{c}
X^Ty \\
Z^Ty
\end{array}
 \right]
$$

Information from data are inserted to the MME. Start with matrix $X$

```{r}
matX <- model.matrix(lm(Weight ~ 0 + BreastCircumference + as.factor(Herd), data = tbl_beef))
attr(matX,"assign") <- NULL
attr(matX,"contrasts") <- NULL
dimnames(matX) <- NULL
```

For matrix $Z$, we have to complete the pedigree

```{r}
vec_fnd_sire <- setdiff(tbl_beef$Sire, tbl_beef$Animal)
vec_fnd_sire <- vec_fnd_sire[!is.na(vec_fnd_sire)]
vec_fnd_dam <- setdiff(tbl_beef$Dam, tbl_beef$Animal)
vec_fnd_dam <- vec_fnd_dam[!is.na(vec_fnd_dam)]
vec_fnd <- c(vec_fnd_sire, vec_fnd_dam)
n_nr_fnd <- length(vec_fnd)
n_nr_obs <- nrow(tbl_beef)
matZ <- cbind(matrix(rep(0, n_nr_fnd*n_nr_obs), nrow = n_nr_obs),
              diag(nrow = n_nr_obs))
```

The inverse numerator relationship $A^{-1}$ is computed based on the pedigree

```{r}
n_nr_ani <- n_nr_fnd + n_nr_obs
ped <- pedigreemm::pedigree(sire = c(rep(NA, n_nr_fnd), tbl_beef$Sire),
                            dam  = c(rep(NA, n_nr_fnd), tbl_beef$Dam),
                            label = c(1:n_nr_ani))
matA_inv <- as.matrix(pedigreemm::getAInv(ped = ped))
```

The coefficient matrix and the right-hand-sides of the MME are

```{r}
lambda <- (1-n_h2_weight) / n_h2_weight
mat_coef <- rbind(cbind(crossprod(matX), crossprod(matX, matZ)), 
                  cbind(crossprod(matZ, matX), crossprod(matZ) + lambda * matA_inv))
vec_y <- tbl_beef$Weight
mat_rhs <- rbind(crossprod(matX, vec_y),
                 crossprod(matZ, vec_y))
mat_sol <- solve(mat_coef, mat_rhs)
mat_sol
```


* Get estimates of fixed effects and predictions of breeding values from solutions of MME

Estimates for the fixed effects are: 

```{r}
n_nr_sol <- nrow(mat_sol)
n_nr_fix <- n_nr_sol - n_nr_ani
mat_sol[1:n_nr_fix,]
```

Predicted breeding values are:

```{r}
mat_sol[(n_nr_fix + 1):n_nr_sol,]
```


* Compute accuracies from inverse of coefficient matrix of MME

The above computed coefficient matrix corresponds to the simplified version. The prediction error variance is obtained from the inverse of the general form of the coefficient matrix. To get from the simplified to the general form, we have to divide by $\sigma_e^2$. 

```{r}
var_p <- var(tbl_beef$Weight)
var_e <- (1-n_h2_weight) * var_p
mat_coef_gen <- mat_coef / var_e
mat_coef_inv <- solve(mat_coef_gen)
mat_C22 <- mat_coef_inv[(n_nr_fix + 1):n_nr_sol,(n_nr_fix + 1):n_nr_sol]
vec_diag_C22 <- diag(mat_C22)
vec_inb <- pedigreemm::inbreeding(ped = ped)
vec_ui <- n_h2_weight * var_p * (1+vec_inb)
vec_Bi <- 1 - vec_diag_C22 / vec_ui
```

The reliabilities (B%) of all animals in the pedigree are 

```{r}
100 * vec_Bi
```

<!-- master-solution-end -->


