---
title: Livestock Breeding and Genomics - `r tools::toTitleCase(params$doctype)` 7
author: Peter von Rohr
date: 2023-10-27
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_md: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Own Performance
```{r prob1-setup, echo=FALSE, message=FALSE, warning=FALSE}
# set.seed(4206)
 n_nr_animal <- 12
 n_weight_mean <- 283
 n_weight_sd <- 3
# vec_weight <- round(rnorm(n_nr_animal, mean = n_weight_mean, n_weight_sd), digits = 0)
# tbl_weight <- tibble::tibble(Animal = c(1:n_nr_animal),
#                              Weight = vec_weight)
# read from file
s_own_perf_url <- "https://charlotte-ngs.github.io/lbgfs2023/data/lbg_ex07_p1_own_perf.csv"
tbl_weight <- readr::read_delim(s_own_perf_url, delim = ",")
# set h2
n_weight_h2 <- 0.2025
```

Given is the dataset with weight observations for `r n_nr_animal` animals. The heritability ($h^2$) for the trait is `r n_weight_h2`. The population mean $\mu$ can assumed to be the mean of the weights in the table below. 

`
```{r tbl-weight, echo=FALSE}
knitr::kable(tbl_weight, booktabs = TRUE, longtable = TRUE, escape = FALSE)
```

The data in the above table can be read from 

```{r, echo=FALSE}
cat(s_own_perf_url, "\n")
```


### Your Tasks

* Compute the breeding values for all animals given in the table above
* Compute the accuracies of the breeding values of all animals shown in the table above.

<!-- your-solution-start
### Your Solution

* Read the data such that you have the phenotypes available as a vector or a tibble
* Use the formula for estimating breeding values based on own-performance records
* Show the result for each animal

---  your-solution-end -->


<!-- master-solution-start -->

### Solution
The predicted breeding value $\widehat{u_i}$ of animal $i$ is computed as 

$$\widehat{u_i} = h^2 (y_i - \mu)$$

where $h^2$ is the heritability given in the problem, $y_i$ is observation of animal $i$ and $\mu$ is the poputlation mean. The population mean is to be computed from the mean of the observations. Hence

```{r}
n_mu_weight <- mean(tbl_weight$Weight)
```

The computed mean is used to predict breeding values as follows

```{r}
vec_pred_bv <- n_weight_h2 * (tbl_weight$Weight - n_mu_weight)
tbl_result <- tibble::tibble(Animal = c(1:n_nr_animal),
                             `Predicted Breeding Value` = vec_pred_bv)
```

The accuracy for the predicted breeding value is the same for all animals and corresponds to 

$$r(u,y) = \frac{cov(u,y)}{\sqrt{var(y)  * var(u)}} = \frac{var(u)}{\sqrt{var(y)  * var(u)}} = \sqrt{\frac{var(u)}{var(y)}} = h$$

Adding this to the table of results, we get

```{r}
tbl_result <- dplyr::bind_cols(tbl_result, 
                tibble::tibble(Accuracy = rep(sqrt(n_weight_h2), n_nr_animal)))
```

The results for the predicted breeding values are shown in the following table

```{r}
knitr::kable(tbl_result, booktabs = TRUE, longtable = TRUE)
```


<!-- master-solution-end -->



## Problem 2: Breeding Value Prediction Based on Repeated Observations

```{r SetupEx2, echo=FALSE}
geb_gew <- 52
mu2 <- 170
rep <- 0.65
h2 <- 0.45
y <- 320
mu <- 250
```

Elsa has observations for her birth weight ($`r geb_gew`$ kg) and some more repeated measures for her weight. We assume the heritability to be  $h^2 = `r h2`$. The population mean for the repeated observations of the weight is  $`r mu2`$ kg. The repeatability of the weight measurements is $t = `r rep`$. 

The following tables contains all observed values for the weight. 

```{r WeightTable, echo=FALSE, message=FALSE, warning=FALSE}
nr_measure <- 10
wean_weight <- y
slope <- (wean_weight-geb_gew)/(nr_measure-1)
measure <- c(1:nr_measure)
weight <- round(slope*(measure-1) + geb_gew, digits = 0)
mean_weight <- mean(weight)
# dfWeightTable <- data.frame(Measurement = measure, Weight = weight)
s_rep_meas_url <- "https://charlotte-ngs.github.io/lbgfs2023/data/lbg_ex07_p2_rep_measures.csv"
tblWeightTable <- readr::read_delim(s_rep_meas_url, delim = ",")
knitr::kable(tblWeightTable, booktabs = TRUE, longtable = TRUE)
```

The data in the above table can be read from 

```{r, echo=FALSE}
cat(s_rep_meas_url, "\n")
```


a) Predict the breeding value for Elsa based on the repeated weight records.
b) What is the reliability for the predicted breeding value from 2a)?
c) Compare the reliability from 2b) with the reliability that would result from a prediction of breeding values based on own performance.


<!-- your-solution-start
### Your Solution

As in Problem 1, you can use the assigned variables. The variables have the following meaning

* nr_measure:   number of measurements of the weight
* h2:           heritability 
* rep:          repeatability
* mean_weight:  mean of the weights
* mu2:          population mean

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

a) The predicted breeding value based on repeated records is

```{r ResultHatARepMeas, echo=FALSE}
hat_a_rep_meas <- round((nr_measure * h2)/(1+(nr_measure - 1)*rep)*(mean_weight - mu2), digits = 2)
```

$$\hat{u}_i = \frac{nh^2}{1+(n-1)t}(\bar{y}_i - \mu) 
            = \frac{`r nr_measure`*`r h2`}{1+(`r nr_measure-1`*`r rep`)}(`r mean_weight` - `r mu2`)
            = `r hat_a_rep_meas`$$


b) The reliability for the predicted breeding value from 2a) is

```{r relreprecord, echo=FALSE}
rel_rep_rec <- (nr_measure * h2)/(1+(nr_measure - 1)*rep)
```

$$B = r_{u,\bar{y}}^2  = b = \frac{nh^2}{1+(n-1)t} 
                 = \frac{`r nr_measure`*`r h2`}{1+(`r nr_measure-1`*`r rep`)}
                 = `r round(rel_rep_rec, digits = 2)`$$
                 
c) The reliability of the predicted breeding values based on repeated records is larger than the reliability of the prediction based on one record. The relation between the two reliabilities is 

$$ \frac{r_{u,\bar{y}}^2}{r_{u,y}^2} = \frac{n}{1+(n-1)t} = \frac{`r nr_measure`}{1+(`r nr_measure-1`*`r rep`)}
  = `r round(nr_measure/(1+(nr_measure-1)*rep), digits = 2)`$$


<!-- master-solution-end -->



## Problem 3: Predict Breeding Values Based on Progeny Records

```{r prob3-def, echo=FALSE}
n_nr_progeny <- 5
```

A few years later Elsa was the dam of `r n_nr_progeny` offspring. Each of the offspring has a record for weaning weight. Predict the breeding value of Elsa for weaning weight based on the offpsring records listed in the following table.

```{r tab-prog-bw, echo=FALSE, message=FALSE, warning=FALSE}
# vec_ww_prog <- round(y + rnorm(n_nr_progeny, mean = 0, sd = 2), digits = 0)
# tbl_prog_ww <- tibble::tibble(Offspring = c(1:n_nr_progeny),
#                               `Weaning Weight` = vec_ww_prog)
s_prog_rec_url <- "https://charlotte-ngs.github.io/lbgfs2023/data/lbg_ex07_p3_prog_rec.csv"
tbl_prog_ww <- readr::read_delim(s_prog_rec_url, delim = ",")
knitr::kable(tbl_prog_ww, booktabs = TRUE, longtable = TRUE)
```

The data in the above table can be read from 

```{r, echo=FALSE}
cat(s_prog_rec_url, "\n")
```

The mean and the heritability can be taken the same as in Problems 1 and 2 resulting in 

$h^2 = `r h2`$ and $\mu = `r mu`$



<!-- your-solution-start
### Your Solution
This problem can be solved analogously to the previous problems, just taking into account that the information now comes from offspring.

---  your-solution-end -->


<!-- master-solution-start -->

### Solution
The predicted breeding value based on progeny records is defined as

\begin{equation}
\hat{u_i} = b * (\bar{y_i} - \mu)
\end{equation}

where $\bar{y_i}$ corresponds to the mean of the progeny records for animal $i$, and $b$ is the regression coefficient which can be shown to be 

$$b = \frac{2n}{n + k}$$

where $n$ is the number of offspring and $k$ corresponds to 

$$k = \frac{4-h^2}{h^2}$$

Inserting the numbers given in the problem task results in

```{r compute-k-prob3, echo=FALSE}
k_prob3 <- (4-h2)/h2
k_prob3_rounded <- round(k_prob3, digits = 2)
```

$$k = \frac{4 - `r h2`}{`r h2`} = `r k_prob3_rounded`$$

Using the computed value of $k$ allows to get the regression coefficient $b$. 

```{r compute-b-prob3, echo=FALSE}
b_prob3 <- 2 * n_nr_progeny /(n_nr_progeny + k_prob3)
b_prob3_rounded <- round(b_prob3, digits = 2)
```

$$b = \frac{2 * `r n_nr_progeny`}{`r n_nr_progeny` + `r k_prob3_rounded`} = `r b_prob3_rounded`$$

The predicted breeding value based on progeny records corresponds to 

```{r compute-hatui-prob3, echo=FALSE}
baryi_prob3 <- mean(tbl_prog_ww$`Weaning Weight`)
baryi_prob3_rounded <- round(baryi_prob3, digits = 2)
hatui_prob3 <- b_prob3 * (baryi_prob3 - mu)
hatui_prob3_rounded <- round(hatui_prob3, digits = 2)
```

$$\hat{u_i} = `r b_prob3_rounded` * (`r baryi_prob3_rounded` - `r mu`) = `r hatui_prob3_rounded`$$


<!-- master-solution-end -->
 
