---
title: Livestock Breeding and Genomics - `r tools::toTitleCase(params$doctype)` 9
author: Peter von Rohr
date: 2023-11-10
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_md: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Numerator Relationship Matrix
```{r lbg-ex07-p01-setup, echo=FALSE, message=FALSE, warning=FALSE}
s_ped_data <- "https://charlotte-ngs.github.io/lbgfs2023/data/ped_num_rel_mat.csv"
```

The following pedigree is given

```{r, echo=FALSE}
nr_animal <- 6
tbl_pedigree <- readr::read_delim(file = s_ped_data, 
                                  col_types = readr::cols(Calf = readr::col_integer(),
                                                          Sire = readr::col_integer(),
                                                          Dam = readr::col_integer()))
knitr::kable(tbl_pedigree, booktabs = TRUE, longtable = TRUE)
```

The pedigree can be read from the file

```{r, echo=FALSE}
cat(s_ped_data, "\n")
```

Compute the numerator relationship matrix $A$ for the given pedigree. Recall from the course notes that elements of matrix $A$ are computed differently for elements on the diagonal and for off-diagonal elements. In summary, we compute

* diagonal element $(A)_{ii}$ as $(A)_{ii} = 1+F_i$ where $F_i = 0.5 * (A)_{sd}$ where $s$ and $d$ are parents of $i$. 
* off-diagonal element $(A)_{ki}$ as $(A)_{ki} = 0.5 * \left[(A)_{ks} + (A)_{kd}\right]$ where $s$ and $d$ are parents of $i$


### Task
Use two nested loops over the rows and the columns of matrix $A$ to compute all the elements of matrix $A$ using the formulas given above.

<!-- your-solution-start
### Your Solution

* Read the pedigree from the given file
* Add parents that do not appear as animals to the column of animals
* Create an empty square matrix $A$ with as many rows and columns as we have animals in the pedigree.
* Use the above formula to compute the diagonal element of the first row
* Create a first loop to compute all the off-diagonal elements of the first row
* Create a loop to repeat the computations over all rows.

---  your-solution-end -->


<!-- master-solution-start -->

### Solution
Pedigree is read from the given file

```{r}
tbl_ped <- readr::read_delim(file = s_ped_data, 
                             delim = ";",
                             col_types = readr::cols(
                               Calf = readr::col_integer(),
                               Sire = readr::col_integer(),
                               Dam  = readr::col_integer()
                             ))
tbl_ped
```

Find animals that appear only as parents, start with sires

```{r}
vec_founder_sire_idx <- sapply(tbl_ped$Sire[!is.na(tbl_ped$Sire)], 
                               function(x) !is.element(x, tbl_ped$Calf), 
                               USE.NAMES = FALSE)
vec_founder_sire <- unique(tbl_ped$Sire[vec_founder_sire_idx])
vec_founder_sire
```

The same for the dams

```{r}
vec_founder_dam_idx <- sapply(tbl_ped$Dam[!is.na(tbl_ped$Dam)], 
                              function(x) !is.element(x, tbl_ped$Calf), 
                              USE.NAMES = FALSE)
vec_founder_dam <- unique(tbl_ped$Dam[vec_founder_dam_idx])
vec_founder_dam
```

Combining them into one vector and sorting them 

```{r}
vec_founder <- c(vec_founder_sire, vec_founder_dam)
vec_founder <- vec_founder[order(vec_founder)]
vec_founder
```

For each of the founders, a new row is added to the top of the pedigree

```{r}
tbl_ped_numrelmat <- NULL
for (n_founder_idx in vec_founder){
  tbl_ped_numrelmat <- dplyr::bind_rows(tbl_ped_numrelmat, 
                                        tibble::tibble(Calf = n_founder_idx, 
                                                       Sire = NA,
                                                       Dam = NA))
}
tbl_ped_numrelmat <- dplyr::bind_rows(tbl_ped_numrelmat, tbl_ped)
n_ani_ped <- nrow(tbl_ped_numrelmat)
tbl_ped_numrelmat
```

An empty matrix $A$ is initialized

```{r}
mat_A <- matrix(nrow = n_ani_ped, ncol = n_ani_ped)
mat_A
```

Start the computation with the diagonal element for animal $1$

```{r}
i <- 1
s <- tbl_ped_numrelmat$Sire[i]
d <- tbl_ped_numrelmat$Dam[i]
F1 <- ifelse(is.na(s) || is.na(d), 0, 0.5*mat_A[s,d])
mat_A[1,1] <- 1 + F1
mat_A
```

The off-diagonal elements of the first row are computed in using the following loop

```{r}
i <- 1
for (col_idx in 2:n_ani_ped){
  u <- tbl_ped_numrelmat$Sire[col_idx]
  mat_elem_u <- ifelse(is.na(u), 0, mat_A[i,u])
  v <- tbl_ped_numrelmat$Dam[col_idx]
  mat_elem_v <- ifelse(is.na(v), 0, mat_A[i,v])
  mat_A[i,col_idx] <- 0.5 * (mat_elem_u + mat_elem_v)
  mat_A[col_idx,i] <- mat_A[i,col_idx]
}
mat_A
```

This should now be done for all rows. Hence, we create a second loop that runs form the first row until the last row.

```{r}
for (row_idx in 1:n_ani_ped){
  # diagonal element
  s <- tbl_ped_numrelmat$Sire[row_idx]
  d <- tbl_ped_numrelmat$Dam[row_idx]
  Fi <- ifelse(is.na(s) || is.na(d), 0, 0.5*mat_A[s,d])
  mat_A[row_idx,row_idx] <- 1 + Fi

  # off-diagonal
  if (row_idx < n_ani_ped){
    for (col_idx in (row_idx+1):n_ani_ped){
      u <- tbl_ped_numrelmat$Sire[col_idx]
      mat_elem_u <- ifelse(is.na(u), 0, mat_A[row_idx,u])
      v <- tbl_ped_numrelmat$Dam[col_idx]
      mat_elem_v <- ifelse(is.na(v), 0, mat_A[row_idx,v])
      mat_A[row_idx,col_idx] <- 0.5 * (mat_elem_u + mat_elem_v)
      mat_A[col_idx,row_idx] <- mat_A[row_idx,col_idx]
    }
  }
}
mat_A
```


<!-- master-solution-end -->


## Problem 2: Verification
Use the function `pedigreemm::getA()` from package `pedigreemm` to verify your result from problem 1.

<!-- your-solution-start
### Your Solution

* Define the pedigree
* Compute the relationship matrix with `getA()`

---  your-solution-end -->


<!-- master-solution-start -->

### Solution
The pedigree is defined by

```{r}
ped <- pedigreemm::pedigree(sire = tbl_ped_numrelmat$Sire,
                            dam = tbl_ped_numrelmat$Dam,
                            label = as.character(tbl_ped_numrelmat$Calf))

```

The numerator relationship matrix is computed by

```{r, message=FALSE, warning=FALSE}
mat_num_relmat <- pedigreemm::getA(ped = ped)
mat_num_relmat
```

Check whether matrices are identical

```{r}
mat_A - as.matrix(mat_num_relmat)
```

<!-- master-solution-end -->


## Problem 3: Functions in R
Computations such as the computation of the diagonal elements or such as the one of the off-diagonal elements can be isolated and factored out in a important programming construct which is called `function`. A function takes a set of input parameter and transforms them into a result which is returned. For our example of the numerator relationship matrix two functions can be constructed according to the following template

```{r}
compute_square <- function(pn_number){
  square_result <- pn_number*pn_number
  return(square_result)
}
```

The function can be used by function calls which take a given input and return a result

```{r}
compute_square(pn_number = 3)
compute_square(1:10)
```


### Task
Use the above template to construct a function and factor out the computations of the diagonal elements and of the off-diagonal elements into two separate functions.


<!-- your-solution-start
### Your Solution

* Write a function to compute a diagonal element
* Write a function to compute an off-diagonal element
* Use the functions to compute all elements of the numerator relationship matrix

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

The function to compute the diagonal elements can be defined as follows

```{r}
compute_diag_elem <- function(pmat_numrel, ptbl_ped, pn_ani_idx){
   s <- ptbl_ped$Sire[pn_ani_idx]
   d <- ptbl_ped$Dam[pn_ani_idx]
   Fi <- ifelse(is.na(s) || is.na(d), 0, 0.5*pmat_numrel[s,d])
   result_diag_elem <- 1 + Fi
   return(result_diag_elem)
}
```

The function to compute an offdiagonal element is defined below

```{r}
compute_off_diag_elem <- function(pmat_numrel, ptbl_ped, pn_row_idx, pn_col_idx){
  u <- ptbl_ped$Sire[pn_col_idx]
  mat_elem_u <- ifelse(is.na(u), 0, pmat_numrel[pn_row_idx,u])
  v <- ptbl_ped$Dam[pn_col_idx]
  mat_elem_v <- ifelse(is.na(v), 0, pmat_numrel[pn_row_idx,v])
  result_off_diag_elem <- 0.5 * (mat_elem_u + mat_elem_v) 
  return(result_off_diag_elem)
}
```

The functions are used to compute all elements of the matrix as shown below.

```{r}
mat_A <- matrix(nrow = n_ani_ped, ncol = n_ani_ped)
for (row_idx in 1:n_ani_ped){
  # compute diagonal element
  mat_A[row_idx, row_idx] <- compute_diag_elem(pmat_numrel = mat_A, 
                                               ptbl_ped = tbl_ped_numrelmat, 
                                               pn_ani_idx = row_idx)
  # compute off-diagonal elements
  if (row_idx < n_ani_ped){
    for (col_idx in (row_idx+1):n_ani_ped){
      mat_A[row_idx, col_idx] <- compute_off_diag_elem(pmat_numrel = mat_A, 
                                                       ptbl_ped = tbl_ped_numrelmat, 
                                                       pn_row_idx = row_idx, 
                                                       pn_col_idx = col_idx)
      mat_A[col_idx,row_idx] <- mat_A[row_idx, col_idx]
    }
  }
}
mat_A
```

Checking the results

```{r}
mat_A - as.matrix(mat_num_relmat)
```

<!-- master-solution-end -->
 
