---
title: "Decompositions of Numerator Relationship Matrix"
output: html_notebook
---


For a given pedigree, we can compute the numerator relationship matrix as follows

```{r}
ped <- pedigreemm::pedigree(sire = c(rep(NA, 3), 1,3),
                            dam  = c(rep(NA, 3), 2, 2),
                            label = 1:5)
ped
```

The number of animals in the pedigree are given by 

```{r}
(n_nr_ani <- length(ped@label)) 
```


## Numerator Relationship Matrix
The numerator relationship matrix $A$ is obtained as 

```{r}
mat_A <- as.matrix(pedigreemm::getA(ped = ped))
mat_A
```

## Cholesky Factorisation of $A$
The cholesky decomposition is 

```{r}
mat_A_chol <- chol(mat_A)
mat_A_chol
```

## Matrix $D$
The matrix $D$ from the LDL-decomposiiton is obtained as

```{r}
D_mat <- pedigreemm::Dmat(ped = ped)
mat_D <- diag(D_mat, nrow = n_nr_ani)
```

This is obtained from the cholesky decomposition by squaring the diagonal elements 

```{r}
mat_S <- diag(mat_A_chol)
mat_S
```

Comparing that to the result of what we get in `Dmat`

```{r}
mat_S * mat_S - D_mat
```


## Matrix $L$
The matrix L is obtained by setting the diagonal of the cholesky factor to one

```{r}
mat_L <- t(mat_A_chol)
diag(mat_L) <- 1
mat_L
```

## Verification
Verify

```{r}
mat_L %*% mat_D %*% t(mat_L) - mat_A
```


## Inverse Numerator Relationship Matrix
The inverse numerator relationship matrix $A^{-1}$ is computed based on the LDL-decomposition of $A$. Hence from 

$$A = LDL^T$$ 

we get

$$A^{-1} = (L^T)^{-1}D^{-1}L^{-1}$$

This is relevant because $L^{-1}$ and $D^{-1}$ are easier to compute than $A^{-1}$. 


### Matrix $D^{-1}$
Because $D$ is a diagnoal matrix, $D^{-1}$ is a diagonal matrix also, with $i-th$ diagonal element $(D^{-1})_{ii}$ is computed as

$$(D^{-1})_{ii} = (D)_{ii}^{-1} = {1\over (D)_{ii}}$$

For our example from above, we get

```{r}
(mat_D_inv <- diag(1/D_mat, nrow = n_nr_ani))
```


### Matrix $L^{-1}$
The matrix $L^{-1} is obtained from the two compositions of the breeding values with

$$u = P \cdot u + m \Leftrightarrow m = (I-P)u$$

and

$$u = L \cdot m \Leftrightarrow m = L^{-1}u  $$

Setting both of the above expressions for $m$ equal shows that 

$$L^{-1} = I-P$$

Hence, we can initialize the matrix L^{-1} with the identity matrix, then loop through the pedigree and subtract $1/2$ for every parent-offspring position. 

```{r}
mat_L_inv <- diag(nrow = n_nr_ani)
for (i in 1:n_nr_ani){
  cur_ani_id <- as.integer(ped@label[i])
  if (!is.na(ped@sire[i])){
    cur_sire_id <- as.integer(ped@sire[i])
    mat_L_inv[cur_ani_id, cur_sire_id] <- -0.5
  }
  if (!is.na(ped@dam[i])){
    cur_dam_id <- as.integer(ped@dam[i])
    mat_L_inv[cur_ani_id, cur_dam_id] <- -0.5
  }
}
mat_L_inv
```

This can be verified by

```{r}
mat_L_inv - solve(mat_L)
```

### Matrix $(L^T)^{-1}$
The matrix $(L^T)^{-1}$ is the inverse matrix of matrix $L^T$ which means that 

$$L^T \cdot (L^T)^{-1} = I$$

The same relation can be established with the un-transposed version of the matrix $L$

$$L \cdot L^{-1} = L^{-1} \cdot L = I$$

The last equation can be transposed on both sides which gives 

$$(L^{-1} \cdot L)^T = I^T = I$$

Transposition can be brought into the parenthesis which changes the order of the factors leading to

$$L^T \cdot (L^{-1})^T = I$$

Comparing this last equation with the first equation of this subsection implies that 

$$(L^{-1})^T = (L^T)^{-1}$$
which is because of the uniqueness of each inverse matrix. Hence, we can just take the previously computed matrix $L^{-1}$ and transpose it to get the inverse $(L^T)^{-1}$. 

